Tidy data structure to support exploration and modeling of temporal-context data

### 2018-04-26 main.Rmd
added panel data str explanations

However, such richness in temporal data cannot be easily accommodated by currently available data structures in any statistical computing package. In the literature, time series and panel (longitudinal) data are common terms referred to temporal-context data, depending on the research fields. Researchers who are concerned with modelling large $T$ and small $N$ would name as "time series" (serial correlation); those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic). The data format is two-dimensional array, but different modelling focuses lend the data input to column types. A matrix is used to represent multivariate time series, assuming homogeneity (that is, all the columns must be of same type.) and time indices implicitly inferred as attributes (or meta-information).

@@

In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key". In multivariate time series notation as Equation \@ref(eq:mts), $X_{jt}$ represents series $j$, for $j = 1, \dots, p$ and $1 \leq t \leq T$, in the form of a $T \times p$ matrix.

\begin{equation}
  \begin{bmatrix}
  X_{11} & X_{21} & \cdots & X_{p1} \\
  X_{12} & X_{22} & \cdots & X_{p2} \\
  \vdots & \vdots & \ddots & \vdots \\
  X_{1T} & X_{2T} & \cdots & X_{pT}
  \end{bmatrix}
\label{eq:mts}
\end{equation}

>>>>

However, such richness in temporal data cannot be easily accommodated by currently available data structures in any statistical computing package. In the literature, time series and panel (longitudinal) data are common terms referred to temporal-context data, depending on the research fields. Researchers who are concerned with modelling large $T$ and small $N$ would name as "time series" (serial correlation); those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic). The data format is two-dimensional array, but different modelling focuses lend the data input to column types. A matrix is used to represent multivariate time series where each row represents observations measured at a time point and each column represents a series. This matrix representation requires homogeneity (that is, all the columns must be of same type.), time indices implicitly inferred as attributes or meta-information, series of same time length, and explicit missingness. By contrast, panel data are organised in two-dimensional array of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column, due to commonly arisen unbalanced panels. This specification requires explicitly declared panel variable and index, which has been implemented in Stata's **tsset** and R package **plm**. This data organisation appears more advantageous than matrix in supporting explicit time index, multiple subjects of different time lengths.

@@

In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key".
